Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-06-15 edubots_surveys.Rnw
Updated citations in State of the Art Section
All citations updated ref #6

<<setup, echo=FALSE,message=FALSE>>=
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(ggthemes)
library(tidyverse)

@@

survey.pilot %>%  mutate(Messaging.Organization =str_split(Messaging.Organization, ", ")) %>% unnest(cols = c(Messaging.Organization)) -> survey.pilot.MessagingOrganization

survey.pilot %>% group_by( PostCOVID.changes ) %>% filter( n() > 1 ) -> survey.pilot.postCOVID

survey.pilot %>%  mutate(ChatbotType = ifelse(Agenda == "No","","Agenda,")) %>%
  mutate(ChatbotType = ifelse(ClassMaterial == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "ClassMaterial,",sep=""))) %>%
  mutate(ChatbotType = ifelse(FAQ == "No",paste(ChatbotType, "",sep=""), paste(ChatbotType, "FAQ,",sep=""))) %>%
  mutate(ChatbotType = ifelse(Grades == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "Grades,",sep=""))) %>%
  mutate(ChatbotType = ifelse(OtherBot == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "OtherBot,",sep=""))) %>%
  mutate(ChatbotType = substr(ChatbotType,1,nchar(ChatbotType)-1))-> survey.pilot.chatbottype
survey.pilot.chatbottype %>%  mutate(ChatbotType =str_split(ChatbotType, ",")) %>% unnest(cols = c(ChatbotType)) -> survey.pilot.ChatbotType

survey.pilot %>%mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline)))->survey.pilot.Discipline

survey.interaction <- read.csv("data/survey-student-interaction-pilot-3-2021-EN.csv")
survey.interaction %>% group_by( Gender ) %>% summarise ( n = n() ) %>% mutate( freqGender = n/sum(n)) -> freqInteraction.Gender
survey.interaction %>% group_by( Age ) %>% summarise ( n = n() ) %>% mutate( freqAge = n/sum(n)) -> freqInteraction.Age
survey.interaction %>% group_by( Experience ) %>% summarise ( n = n() ) %>% mutate( freqExperience = n/sum(n)) -> freqInteraction.Experience
survey.interaction %>% group_by( Vocational ) %>% summarise ( n = n() ) %>% mutate( freqSector = n/sum(n)) -> freqInteraction.Sector
freqInteraction.Sector$Sector <- freqInteraction.Sector$Vocational
freqInteraction.Sector %>% mutate(Sector = ifelse(as.character(Vocational) == "Yes", "Vocational", "University")) -> freqInteraction.Sector
freqInteraction.Sector <- freqInteraction.Sector[-c(1)]
survey.interaction %>% group_by( Discipline ) %>% summarise ( n = n() ) %>% mutate( freqDiscipline = n/sum(n)) -> freqInteraction.Discipline
freqInteraction.Discipline %>%mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) -> freqInteraction.Discipline
freq.Discipline %>% mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) %>%
                mutate(n = ifelse(as.character(Discipline) =="Engineering", sum(n[as.character(Discipline)=="Engineering" | as.character(Discipline)=="Technology"]), n)) %>%
                mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
                subset(as.character(Discipline) != "Technology")-> freq.Discipline
@

@@

In the framework of the EduBots KA2 European project\footnote{https://www.edubots.eu}, we are mainly concerned with the creation of best practices in the use of chatbots in higher education classrooms. A \textit{chatbot} is a program, sometimes based on Artificial Intelligence techniques, that allows a computer to communicate in a similar way as a human does \cite{Gong2008}, using natural language. Indeed, most of them would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any speaker's question fluently, and even ask other questions to the human. However, in most cases a simple messaging-application attached bot will help the students (and teachers); neither computational intelligence nor comprehension of natural languages are requisites, the only one being the fact that they can be accessed from a synchronous conversational application such as the common instant messaging applications used nowadays.

@@

In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, students suffer from stress when they have to ask a question in front of the entire class. This is why many students prefer to remain silent and ask their question later via email. 
However, this means a waste of time and a delay of hours or even days to get the teacher's answer; it does not scale well with the size of the class.
Chatbot technology has a potential to fill the gap between teacher and students, helping them to solve questions and carry out a dynamic and autonomous learning \cite{Griol2014,kim2020}.
Moreover, using an automated system such as a chatbot allows the teacher to detect the most demanded topics, and on which he or she should consequently place more emphasis in class.

@@

% Maybe move it to a different place later
In the present work we aim to complement the aforementioned study with teacher attitudes; teachers are the ones that actually would need to implement, configure and maybe program chatbots. Thus, we have performed a survey among 300 teachers asking about their attitudes towards the use of bots or chatbots in their classes. We analyse their responses, since we are interested in discussing implications and the impact for the development of future technology-enhanced tools and the Education institutions policies regarding the following research questions:
\begin{itemize}
\item{ RQ1 - Are teachers already using messaging apps in their classes? } %This was the second survey, not the  first survey. I wonder if maybe we could invert the presentation order - JJ
\item{ RQ2 - Which kind of chatbots would teachers find useful in their classes? }
\item{ RQ3 - Which kind of interaction do teachers prefer with their students? }

@@

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant messaging (MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook Messenger stems from their simplicity, ease of use and multi-modality (i.e. video, audio, text) \cite{tang2017mobile}.  Using these tools in higher education can facilitate the delivery of personalised learning that occurs anytime anywhere, promote collaborative learning experiences and group discussions \cite{panah2020study}. WhatsApp is the most popular MIM platform used by educators to give assignments' feedback to students, support course discussions and provide learning resources in informal learning settings \cite{panah2020study}.  The use of WhatsApp in higher education could enhance social presence \cite{tang2017mobile} and foster trust relationships between educators and students embedded in the social learning process \cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due to the blurring of boundaries between academic and private life. This can lead to technostress \cite{gachago2015crossing}, difficulty in managing responsibilities especially among mature students and lack of privacy \cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave groups at any time can hinder their learning and undermine educators' efforts \cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and norms for these MIM groups in order to maintain the safety of these online communities for students (Abbas et al., 2021b) % This one is missing from the biblio
without affecting students' ownership and control that is vital to advance their learning \cite{gachago2015crossing}.

% The Use of Chatbots in Higher Education
The use of conversational agents or chatbots in higher education is still at its infancy cite{yang2019opportunities}. Nevertheless, recent studies examining their positive impact on students' academic performance \cite{perez2020rediscovering} and engagement \cite{differ.chat}(Abbas et al., 2021b) have led to a growing interest in using this technology in classes. Indeed, the use of chatbots in collecting course feedback from students in higher education improved students' response quality and boosted their enjoyment levels (Abbas, Pickard, Atwell and Walker, 2021a).  According to \cite{Roblyer2010}, using either tools such as mobile devices or teaching strategies based on gamification \cite{Yildirim2017} can improve student motivation. In this sense, \cite{Pimmer2019} carried out a series of experiments with technological tools whose results showed positive perception and acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies were undertaken. For instance, an evaluation review study presented by \cite{Smutny2020} examined 47 educational chatbots implemented in Facebook Messenger with the focus to identify characteristics and quality metric such as language, subject matter and platform. Another study aimed to categorise educational chatbots according to their purpose into service-oriented and teaching-oriented \cite{perez2020rediscovering}. The first category includes those that provide service support such as the chatbot Ask Holly (University of Durham, 2021) % We probably need a reference for this - JJ
and Dina \cite{santoso2018dinus}, both chatbots respond to students' questions about enrolment and registration. Ask L.U. (Lancaster University, 2019 % Another reference
) answers students' frequently asked questions about timetables, grades, tutors and societies. LISA \cite{dibitonto2018chatbot} and Differ (Differ, 2021) facilitate breaking the ice between new students by introducing them to each other.  \cite{ranoliya2017chatbot} proposed a generic chatbot for university students that is able to answer their frequently asked questions; for some time, the University of Granada had ELVIRA as a chatbot in their main page \cite{MOREO20129}. Besides being able to answer pre-established (frequent) questions, it did so from the main University of Granada using an inset character, voiced by a real person, who lip-synched the answers. The emphasis in this case, however, was on the authenticity of the synchronization more than in making the update of questions and answers simpler, or more interactive, using actual messaging platforms. In most cases, using the inset search engine was able to give you more up-to-date and accurate answers.

On the other hand, teaching-oriented chatbots are more sophisticated as they set personalised learning outcomes and monitor learning progress. For instance, \cite{fernoagua2018intelligent} reported on ''eduAssistant'' a virtual teaching assistant chatbot developed on the Telegram messaging platform. In this study, the Telegram platform was chosen because it is easy to use, students are familiar with its features and it enables students to exchange messages in different formats (text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram could operate on all devices and operating systems. The ''eduAssistant'' chatbot acts as an automatic agent between the teacher-content-student, facilitating real-time feedback loops and providing a personalised learning experience relevant to students' skills and knowledge. Using this chatbot, educators can create interactive instances in their lectures where they pose questions to their students and the chatbot assists students who need further help by giving those more hints and reporting to their educator's dashboard \cite{fernoagua2018intelligent}. This can help educators locate those students that need their attention and send them more educational resources relevant to their academic attainment.

Furthermore, \cite{sjostrom2018designing} proposed a conceptual architecture for teaching-oriented chatbots in higher education.  This conceptual architecture is based on a systematic literature review of previous studies examining the design of chatbots in higher education and a content analysis of student emails and discussion forum posts of four instances of a Java programming course.  The study outlined several design considerations, among them, the authors emphasised the importance of developing chatbots in platforms that students and educators are familiar with and can easily access (i.e. Facebook Messenger) which confirms with \cite{hobert2019you} and \cite{fernoagua2018intelligent} studies. In addition, \cite{sjostrom2018designing} argued that a conceptualision of learners' questions could aid designers in integrating the appropriate types of questions that the chatbots should support for different courses.  \cite{coronado2018cognitive} proposed agents that store learning materials to be provided on demand to students. \cite{crockett2017predicting} reported on tutoring systems, which can perform initial assessments of students' understanding and provide learning material that would advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many studies have focused on the evaluation of technology acceptance and usability \cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain where according to \cite{hobert2019you}, specific pedagogical factors such as learning success or increased motivation are more important.  Therefore, to develop efficient chatbots for higher education, all stakeholders' (i.e. educators, students, institutions, etc) expectations needs to be carefully collected and taken into consideration  \cite{sjostrom2018designing,tsivitanidouusers}.

@@

The initial survey was done in two different forms, one for university and other for tertiary non-university teachers. Questions and responses were the same, except for the type of tertiary degrees that were taught. We got around 300 responses to this one, with two thirds of them coming from university teachers, the rest from high school teachers. In both cases, we used mainly mailing lists, as well as Telegram groups, to spread the link to the form. The university form was spread in Spanish (mainly Andaluc\'{\i}a and Galicia) universities, as well as Costa Rica and Mexico. The tertiary non-university teachers are mainly in Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of 2021, post-pandemic and while, at least in Spain, many universities had mandatory virtual teaching.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for the first survey (\textit{survey1} in Fig. \ref{figure:teachers_population}) and 205 for the second one (\textit{survey2}). The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, in total 61\% are male teachers and 32\% female (approximately 5\% chose not to disclose it). Finally, the responses are equally distributed attending to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

% [pedro]  the way these (5) figures are shown is a bit strange: different sizes of plots...

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<gender, echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Gender, aes(x=Gender, y=freqGender, fill=Gender)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Gender$source<-rep("survey1")
freqInteraction.Gender$source<-rep("survey2")
freqbinded.Gender <- rbind(freq.Gender, freqInteraction.Gender)
freqbinded.Gender$source <- factor(freqbinded.Gender$source) 
ggplot(freqbinded.Gender, aes(x = Gender, y=freqGender, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/gender.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<age,echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Age, aes(x=Age, y=freqAge, fill=Age)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Age$source<-rep("survey1")
freqInteraction.Age$source<-rep("survey2")
freqbinded.Age <- rbind(freq.Age, freqInteraction.Age)
freqbinded.Age$source <- factor(freqbinded.Age$source) 
ggplot(freqbinded.Age, aes(x = Age, y=freqAge, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/age.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<sector,echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Sector, aes(x=Sector, y=freqSector, fill=Sector)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Sector$source<-rep("survey1")
freqInteraction.Sector$source<-rep("survey2")
freqbinded.Sector <- rbind(freq.Sector, freqInteraction.Sector)
freqbinded.Sector$source <- factor(freqbinded.Sector$source) 
ggplot(freqbinded.Sector, aes(x = Sector, y=freqSector, fill = source)) + geom_bar(stat="identity", position = "dodge")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector.pdf}
    \end{minipage}
    \vspace{0.5mm}
    \begin{minipage}[t]{0.38\textwidth}
        \centering
<<experience, echo=FALSE, fig.width=4.5, fig.height=3.475>>=
#ggplot(freq.Experience, aes(x=Experience, y=freqExperience, fill=Experience)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Experience$source<-rep("survey1")
freqInteraction.Experience$source<-rep("survey2")
freqbinded.Experience <- rbind(freq.Experience, freqInteraction.Experience)
freqbinded.Experience$source <- factor(freqbinded.Experience$source) 
ggplot(freqbinded.Experience, aes(x = Experience, y=freqExperience, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@        
%       \includegraphics[width=\textwidth]{figures/experience.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.52\textwidth}
        \centering
<<discipline,echo=FALSE, fig.width=8.2, fig.height=4.45>>=
#ggplot(freq.Discipline, aes(x=Discipline, y=freqDiscipline, fill=Discipline)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Discipline$source<-rep("survey1")
freqInteraction.Discipline$source<-rep("survey2")
freqbinded.Discipline <- rbind(freq.Discipline, freqInteraction.Discipline)
freqbinded.Discipline$source <- factor(freqbinded.Discipline$source) 
freqbinded.Discipline$Discipline <- factor(freqbinded.Discipline$Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other"))
ggplot(freqbinded.Discipline, aes(x = Discipline, y=freqDiscipline, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/discipline.pdf}
    \end{minipage}
\end{center}
    \caption{Sample population description. The responses from two surveys were collected amount to 282 for the first survey (\textit{survey1} in red) and 205 for the second (\textit{survey2} in blue). Data are shown by gender, age, sector (university and vocational education), number of years of experience in education, and discipline.}
\label{figure:teachers_population}
\end{figure}

Figure \ref{figure:teachers_population} shows the distribution of age and experience are quite similar. %, and as a matter of fact,
Additionally, as a matter of fact, as Figure \ref{figure:age.vs.exp} shows, age and experience are correlated strongly, with experience being roughly 20-30 years less than age. 
This might indicate that a differential analysis for age and experience will not yield significant differences, except maybe in the 6-15 years experience group, which is spread over all ages. At any rate, experience seems to be the dominating factor here and we'll use experience only unless we detect some significant difference by age.

\begin{figure}[h!tbp]
\begin{center}
<<interaction,message=FALSE,echo=FALSE, fig.height=2.0, fig.width=4.0>>=
library(forcats)
survey.interaction %>%
    mutate(Age = fct_relevel(Age,"25-35", "35-45", "45-55", "> 55")) %>%
    mutate(Experience = fct_relevel(Experience, "0-05 years", "06-15 years", "16-25 years", "> 25 years" )) %>%
    group_by( Age, Experience ) %>% summarise( n = n() ) %>% mutate( freq = n/sum(n)) -> Age.vs.Experience

@@

@@

\begin{figure}[h!tbp]
\centering
<<messaging_organization, echo=FALSE, fig.width=19/2.54, fig.height=5/2.54>>=
survey.pilot.MessagingOrganization <- within(survey.pilot.MessagingOrganization, Messaging.Organization <- factor(Messaging.Organization, levels=names(sort(table(Messaging.Organization), decreasing=TRUE))))
ggplot(survey.pilot.MessagingOrganization, aes( x = Messaging.Organization, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90))+coord_flip()+ theme(axis.title.y=element_blank())
@
%\includegraphics[width=\textwidth]{figures/messaging_organization.pdf}
\caption{Distribution of teachers' preferences for the chat groups with their students: from groups only with their students from a specific course to groups with greater social interaction with all students in their School or Faculty.}
\label{figure:chatbots_messagingorganization}
\end{figure}

>>>>

<<setup, echo=FALSE,message=FALSE>>=
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(ggthemes)

@@

data <- read.csv("data/survey-student-interaction-pilot-3-2021-EN.csv")
@

@@

In the framework of the EduBots KA2 European project, we are mainly concerned with the creation of best practices in the use of chatbots in higher education classrooms. A \textit{chatbot} is a program, sometimes based on Artificial Intelligence techniques, that allows a computer to communicate in a similar way as a human does \cite{Gong2008}, using natural language. Indeed, most of them would be, apparently, close to passing a classic Turing test \cite{moor2003turing}, since they are able to answer almost any speaker’s question fluently, and even ask other questions to the human. However, in most cases a simple messaging-application attached bot will help the students (and teachers); neither computational intelligence nor comprehension of natural languages are requisites, the only one being the fact that they can be accessed from a synchronous conversational application such as the common instant messaging applications used nowadays.

@@

In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, students suffer from stress when they have to ask a question in front of the entire class. This is why many students prefer to remain silent and ask their question later via email. 
However, this means a waste of time and a delay of hours or even days to get the teacher's answer.
Chatbot technology has a potential to fill the gap between teacher and students, helping them to solve questions and carry out a dynamic and autonomous learning \cite{Griol2014,kim2020}.
Moreover, using an automated system such as a chatbot allows the teacher to detect the most demanded topics, and on which he or she should consequently place more emphasis in class.

@@

% Maybe move it to a different place later
In the present work we aim to complement the aforementioned study with teacher attitudes; teachers are the ones that actually would need to implement, configure and maybe program chatbots. Thus, we have performed a survey among 300 teachers asking about their attitudes towards the use of bots or chatbots in their classes. We analyse their responses, since we are interested in discussing implications and the impact for the development of future technology-enhanced tools and the Education institutions policies regarding the following research questions:
\begin{itemize}
\item{ RQ1 - Are teachers already using messaging apps in their classes? } %This was the second survey, not the  first survey. I wonder if maybe we could invert the presentation order - JJ
\item{ RQ2 - Which kind of chatbots would teachers find useful in their classes?? }
\item{ RQ3 - Which kind of interaction do teachers prefer with their students? }

@@

The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free MIM tools/platforms such as WhatsApp, Telegram, WeChat and Facebook Messenger stems from their simplicity, ease of use and multi-modality (i.e. video, audio, text) \cite{tang2017mobile}.  Using these tools in higher education can facilitate the delivery of personalised learning that occurs anytime anywhere, promote collaborative learning experiences and group discussions \cite{panah2020study}. WhatsApp is the most popular MIM platform used by educators to give assignments’ feedback to students, support course discussions and provide learning resources in informal learning settings \cite{panah2020study}.  The use of WhatsApp in higher education could enhance social presence \cite{tang2017mobile} and foster trust relationships between educators and students embedded in the social learning process \cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due to the blurring of boundaries between academic and private life. This can lead to technostress \cite{gachago2015crossing}, difficulty in managing responsibilities especially among mature students and lack of privacy \cite{tang2017mobile}.  Students’ dropout of the MIM groups, as they can leave groups at any time can hinder their learning and undermine educators’ efforts \cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and norms for these MIM groups in order to maintain the safety of these online communities for students \cite{abbas2021onlinechat} without affecting students’ ownership and control that is vital to advance their learning \cite{gachago2015crossing}.

The Use of Chatbots in Higher Education
The use of conversational agents or chatbots in higher education is still at its infancy \cite{yang2019opportunities}. Nevertheless, recent studies examining their positive impact on students’ academic performance \cite{perez2020rediscovering} and engagement \cite{differ.chat, abbas2021onlinechat} have led to a growing interest in using this technology in classes. Indeed, the use of chatbots in collecting course feedback from students in higher education improved students’ response quality and boosted their enjoyment levels \cite{abbas2021Surveys}.  According to Roblyer et al. \cite{Roblyer2010}(2010), using either tools such as mobile devices or teaching strategies based on gamification \cite{Yildirim2017} can improve student motivation. In this sense, Pimmer et al. (2019)\cite{Pimmer2019} carried out a series of experiments with technological tools whose results showed positive perception and acceptance of the use of technology in teaching and learning.

Several higher education chatbots’ evaluation studies were undertaken. For instance, an evaluation review study presented by Smutny and Schreiberova (2020)\cite{Smutny2020} examined 47 educational chatbots implemented in Facebook Messenger with the focus to identify characteristics and quality metric such as language, subject matter and platform. Another study aimed to categorise educational chatbots according to their purpose into service-oriented and teaching-oriented \cite{perez2020rediscovering}. The first category includes those that provide service support such as the chatbot Ask Holly \cite{meetholly} and Dina \cite{santoso2018dinus}, both chatbots respond to students’ questions about enrolment and registration. Ask L.U. \cite{lancasterU} answers students’ frequently asked questions about timetables, grades, tutors and societies. LISA \cite{dibitonto2018chatbot} and Differ \cite{differchat} facilitate breaking the ice between new students by introducing them to each other.  Ranoliya et al. (2017)\cite{ranoliya2017chatbot} proposed a generic chatbot for university students that is able to answer their frequently asked questions.

On the other hand, teaching-oriented chatbots are more sophisticated as they set personalised learning outcomes and monitor learning progress. For instance, Fernoagă, Stelea, Gavrilă, and Sandu (2018) \cite{ernoagua2018intelligent} reported on ‘eduAssistant’ a virtual teaching assistant chatobt developed on the Telegram messaging platform. In this study, the Telegram platform was chosen because it is easy to use, students are familiar with its features and it enables students to exchange messages in different formats (text, audio and video) \cite{ernoagua2018intelligent}. In addition, Telegram could operate on all devices and operating systems. The ‘eduAssistant’ chatbot acts as an automatic agent between the teacher-content-student, facilitating real-time feedback loops and providing a personalised learning experience relevant to students’ skills and knowledge. Using this chatbot, educators can create interactive instances in their lectures where they pose questions to their students and the chatbot assists students who need further help by giving those more hints and reporting to their educator’s dashboard \cite{ernoagua2018intelligent}. This can help educators locate those students that need their attention and send them more educational resources relevant to their academic attainment.  
Furthermore, Sjöström et al. (2018) \cite{Sjostrom2018} proposed a conceptual architecture for teaching-oriented chatbots in higher education.  This conceptual architecture is based on a systematic literature review of previous studies examining the design of chatbots in higher education and a content analysis of student emails and discussion forum posts of four instances of a Java programming course.  The study outlined several design considerations, among them, the authors emphasised the importance of developing chatbots in platforms that students and educators are familiar with and can easily access (i.e. Facebook Messenger) which confirms with Hobert’s (2019) \cite{hobert2019you} and Fernoagă, Stelea, Gavrilă, and Sandu’s (2018) \cite{fernoagua2018intelligent}  studies. In addition, Sjöström et al. \cite{Sjostrom2018} argued that a conceptualision of learners’ questions could aid designers in integrating the appropriate types of questions that the chatbots should support for different courses.  Coronado et al. (2018) \cite{coronado2018cognitive} proposed agents that store learning materials to be provided on demand to students. Crockett et al. (2017) \cite{crockett2017predicting} reported on tutoring systems, which can perform initial assessments of students’ understanding and provid learning material that would advance their understanding to the next level.
Regarding the factors for the adoption of chatbots in higher education, many studies have focused on the evaluation of technology acceptance and usability \cite {Roblyer2010, Pimmer2019}. However, higher education is a special domain where according to Hobert (2019) \cite{hobert2019you}, specific pedagogical factors such as learning success or increased motivation are more important.  Therefore, to develop efficient chatbots for higher education, all stakeholders’ (i.e. educators, students, institutions, etc) expectations needs to be carefully collected and taken into consideration \cite{sjostrom2018designing, tsivitanidouusers}.

@@

The initial survey was done in two different forms, one for university and other for tertiary non-university teachers. Questions and responses were the same, except for the type of tertiary degrees that were taught. We got around 300 responses to this one, with two thirds of them coming from university teachers, the rest from high school teachers. In both cases, we used mainly mailing lists, as well as Telegram groups, to spread the link to the form. The university form was spread in Spanish (mainly Andalucía and Galicia) universities, as well as Costa Rica and Mexico. The tertiary non-university teachers are mainly in Andalucía. Dissemination and answers took place in the first quarter of 2021, post-pandemic and while, at least in Spain, many universities had mandatory virtual teaching.

@@

% Describe the survey and population
We collected responses from a 282 teachers from Spain and Spanish Speaking countries. The sample includes teachers from university (68\%) and vocational Education (32\%). Regarding gender, 63\% are male teachers and 35\% female (approximately 2\% chose not to disclose it). Finally, the responses are equally distributed attending to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 27\% for 6-15 years of experience, 30\% for 16-25, and 19\% for teachers with more than 25 years of experience.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.3\textwidth}
	    \centering
<<gender, echo=FALSE>>=
ggplot(freq.Gender, aes(x=Gender, y=freqGender, fill=Gender)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@

% 	    \includegraphics[width=\textwidth]{figures/gender.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
	    \centering
<<age,echo=FALSE>>=
ggplot(freq.Age, aes(x=Age, y=freqAge, fill=Age)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@

% 	    \includegraphics[width=\textwidth]{figures/age.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
	    \centering
<<sector, echo=FALSE>>=
ggplot(freq.Experience, aes(x=Experience, y=freqExperience, fill=Experience)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@

% 	    \includegraphics[width=\textwidth]{figures/sector.pdf}
    \end{minipage}

    \vspace{0.5mm}
    \begin{minipage}[t]{0.3\textwidth}
	    \centering
<<experience,echo=FALSE>>=
ggplot(freq.Sector, aes(x=Sector, y=freqSector, fill=Sector)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@

% 	    \includegraphics[width=\textwidth]{figures/experience.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.55\textwidth}
	    \centering
<<discipline,echo=FALSE>>=
ggplot(freq.Discipline, aes(x=Discipline, y=freqDiscipline, fill=Discipline)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
@

% 	    \includegraphics[width=\textwidth]{figures/discipline.pdf}
    \end{minipage}
\end{center}
    \caption{Sample population description. A total 282 teachers shown by gender, age, sector (university and vocational education), number of years of experience in education, and discipline.}
\label{figure:teachers_population}
\end{figure}
%
\begin{figure}[h!tbp]
\begin{center}
<<interaction,message=FALSE,echo=FALSE>>=
library(forcats)
data %>%
    mutate(Age = fct_relevel(Age,"25-35", "35-45", "45-55", "> 55")) %>%
    mutate(Experience = fct_relevel(Experience, "0-05 years", "06-15 years", "16-25 years", "> 25 years" )) %>%
    group_by( Age, Experience ) %>% summarise( n = n() ) %>% mutate( freq = n/sum(n)) -> Age.vs.Experience

@@

Figure \ref{figure:teachers_population} shows the distribution of age and experience are quite similar, and as a matter of fact, as Figure \ref{figure:age.vs.exp} age and experience are correlated strongly, with experience being roughly 20-30 years less than age. This might indicate that a differential analysis for age and experience will not yield significant differences, except maybe in the 6-15 years experience group, which is spread over all ages. At any rate, experience seems to be the dominating factor here and we'll use experience only unless we detect some significant difference by age.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{figures/messaging_organization.pdf}
\caption{Distribution of teachers' preferences for the chat groups with their students: from groups only with their students from a specific course to groups with greater social interaction with all students in their School or Faculty.}
\label{figure:chatbots_messagingorganization}
\end{figure}
